Machine-learned embedded atom method

### 2019-10-16 doc/papers/ml-eam/manuscript.tex
The ML-EAM manuscript is updated.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EAM - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}[h!]
\centering
\includegraphics[scale=0.57]{figures/Ni_eam.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EOS
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.65]{figures/Ni_all_eos.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure}

We first use the elementary fcc Ni dataset to show our approach. This dataset 
has 461 structures in total. 400 were randomly selected to form the training 
subset. In this experiment, $\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and 
$\chi_{\mathrm{elastic}}$ are set to 1, 3 and 0.05 respectively. In this 
section, SNAP shall refer to the elementary SNAP Ni model. Table 
\ref{table:elementary_Ni_mae} shows the MAEs of those models. The original EAM 
model already has an accepetable performace on this dataset. Our machine 
learning approach can further reduce the energy MAE to 4.1 meV/atom. All of 
these models have similar force prediction performance.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni - MAE
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lccc}
\hline
    & SNAP & EAM & ML-EAM \\
\hline
Energy MAE (meV/atom) & 1.2 & 10.6 & 4.1 \\
Force MAE (eV/\AA) & 0.05 & 0.06 & 0.05 \\
\hline
\end{tabular}
\caption{\label{table:elementary_Ni_mae} Comparion of the energy and force 
accuracy of the elementary SNAP Ni mode, the original EAM and the machine 
learned EAM.
}
\end{table}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni - Material Properties
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}[h]
\centering
\begin{tabular}{lcccccc}
\hline
    & DFT & SNAP & EAM & ML-EAM & Expt. \\
\hline
$T_{m}$(K) & & 1785 & 1520 & 1765 & 1600 & 1728 \\
$c_{11}$ (GPa) & 276 & 276 (0.0\%) & 248 (-10.1\%) & 260 (-5.8\%) & 274 (-0.7\%) & 261 \\
$c_{12}$ (GPa) & 159 & 159 (0.0\%) & 147 (-7.5\%) & 151 (0.0\%) & 163 (2.5\%) & 151 \\
$c_{44}$ (GPa) & 132 & 132 (0.0\%) & 125 (-5.3\%) & 131 (0.0\%) & 131 (-0.8\%) & 132 \\
$E_{v}$ (eV) & 1.46 & 1.68 (15.1\%) & 1.68 (15.1\%) & 1.16 (-20.5\%) & 1.71 (17.1\%) & 1.54–1.80 \\
$E_{m}$ (eV) & 1.12 & 1.07 (-4.5\%) & 0.90 (-19.6\%) & 1.46 (30.5\%) & 0.87 (-22.3\%) & 1.01–1.48 \\
$E_{a} = E_{v} + E_{m} $ (eV) & 2.58 & 2.75 (6.6\%) & 2.58 (0.0\%) & 2.62 (1.6\%) & 2.58 (0.0\%) & 2.77–2.95 \\
\hline
\end{tabular}
\caption{\label{table:elementary_Ni} Comparison of the calculated and 
experimental material properties of elementary Ni. $T_{m}$ is the melting 
point. $c_{ij}$ represents elastic constants. $E_{v}$ and $E_{m}$ are vacancy 
formation energy and migration energy, respectively. SNAP refers to the 
elementary Ni SNAP model.
}
\end{table*}

@@

>>>>

The melting points $T_{m}$ of bcc Mo and fcc Ni were measured with the 
solid-liquid coexistence approach. For fcc Ni, $30 \times 10 \times 10$ 
supercells were used and for bcc the supercells were $30 \times 15 \times 15$. 
The time step was set to 1 fs. The simulation for each temperature was carried 
out for 300 ps.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EOS
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[htp]
\centering
\includegraphics[scale=0.65]{figures/Ni_all_eos.png}
\caption{\label{fig:Ni_eam_eos} The energy-volume curves of the original EAM, 
SNAP, ML-EAM and DFT for the fcc Ni.}
\end{figure}

We first use the elementary fcc Ni dataset to show our approach. This dataset 
has 461 structures in total. 400 were randomly selected to form the training 
subset. In this experiment, $\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and 
$\chi_{\mathrm{elastic}}$ were set to 1, 3 and 0.05 respectively. The batch size
was 25. Only the fcc Ni was included in $\Omega$ (Equation \ref{eq:rose_loss} 
and \ref{eq:cijkl_loss}).

Table \ref{table:elementary_Ni_mae} shows the MAEs of those models. The original 
EAM model already has an accepetable performace on this fcc dataset. Its energy 
MAE is 10.6 meV/atom and the force MAE is 0.06 eV/\AA. The MEAM performs 
slightly worse on this dataset with an energy MAE of 17.6 meV/atom and force MAE
of 0.08 eV/\AA. Our machine learning approach can further reduce the energy MAE 
to 4.1 meV/atom and the force MAE to 0.05 eV/\AA, which are much closer to the 
elementary SNAP Ni model.

Figure \ref{fig:Ni_eam_eos} plots the equation of state curves obtained with the 
DFT, SNAP, EAM and ML-EAM models. The whole fitting region ranges from 84\% to 
118\% of the equilibrium volume. The original EAM significantly underestimates 
the energy at both tensile and compress strains. However, our machine learning 
approach together with the Rose constraint can fix this problem. The 
energy-volume curve of ML-EAM overalps with the DFT curve very well for the 
entire fitting region.

Table \ref{table:elementary_Ni} summerizes the predicted material properties of 
fcc Ni, including the melting point, the elastic constants, the vacancy 
formation energy $E_{v}$, the migration energy $E_{m}$ and the activation energy 
$E_{a}$. The original EAM already performs well on these properties. ML-EAM 
further improves the elastic constants. The errors were significantly reduced 
to around 1-2\%.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni - MAE
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lcccc}
\hline
 MAE & SNAP & EAM & MEAM & ML-EAM \\
\hline
Energy (meV/atom) & 1.2 & 10.6 & 17.8 & 4.1 \\
Force (eV/\AA) & 0.05 & 0.06 & 0.08 & 0.05 \\
\hline
\end{tabular}
\caption{\label{table:elementary_Ni_mae} Comparion of the energy and force 
accuracy of the elementary SNAP Ni model, the original EAM, the MEAM and our 
ML-EAM.
}
\end{table}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni - Material Properties
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}[htp]
\centering
\begin{tabular}{lcccccc}
\hline
    & DFT & SNAP & EAM & MEAM & ML-EAM & Exp. \\
\hline
$T_{m}$(K) & & 1785 & 1520 & 1765 & 1520 & 1728 \\
$c_{11}$ (GPa) & 276 & 276 (0.0\%) & 248 (-10.1\%) & 260 (-5.8\%) & 274 (-0.7\%) & 261 \\
$c_{12}$ (GPa) & 159 & 159 (0.0\%) & 147 (-7.5\%) & 151 (0.0\%) & 163 (2.5\%) & 151 \\
$c_{44}$ (GPa) & 132 & 132 (0.0\%) & 125 (-5.3\%) & 131 (0.0\%) & 131 (-0.8\%) & 132 \\
$E_{v}$ (eV) & 1.46 & 1.68 (15.1\%) & 1.68 (15.1\%) & 1.16 (-20.5\%) & 1.71 (17.1\%) & 1.54–1.80 \\
$E_{m}$ (eV) & 1.12 & 1.07 (-4.5\%) & 0.90 (-19.6\%) & 1.46 (30.5\%) & 0.87 (-22.3\%) & 1.01–1.48 \\
$E_{a} = E_{v} + E_{m} $ (eV) & 2.58 & 2.75 (6.6\%) & 2.58 (0.0\%) & 2.62 (1.6\%) & 2.58 (0.0\%) & 2.77–2.95 \\
\hline
\end{tabular}
\caption{\label{table:elementary_Ni} Comparison of the calculated and 
experimental material properties of elementary fcc Ni. $T_{m}$ is the melting 
point. $c_{ij}$ represents elastic constants. $E_{v}$ and $E_{m}$ are vacancy 
formation energy and migration energy, respectively. SNAP refers to the 
elementary Ni SNAP model.
}
\end{table*}

Figure \ref{fig:Ni_eam} compares the $rho(r)$, $F(\rho)$ and $\phi(r)$ functions
of ML-EAM with original Zjw04 EAM. The pairwise potential $\phi(r)$ almost 
remains the same while the embedding function $F(\rho)$ and the electron density 
function $\rho(r)$ changes significantly. Both $\rho(r)$ and $\phi(r)$ converge
to zero at around $r=4.8$ \AA, which is rougly 2 times of the equilibrium 
spacing between nearest neighbors.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EAM - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}[htp]
\centering
\includegraphics[scale=0.57]{figures/Ni_eam.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.B
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Mo}
\label{sec:elementary_Mo}

The second test example is bcc Mo. This dataset has 284 structures. 34 of 
them were used as the testing subset and the rest were used for training. 
The training settings were the same with fcc Ni. An ADP potential was also 
optimized for bcc Mo.

Table \ref{table:elementary_Mo_mae} summerizes the energy and force MAEs on the 
test subset. The original Zjw04 EAM performs significantly worse. The energy

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EAM/ADP - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}[htp]
\centering
\includegraphics[scale=0.6]{figures/Mo_eam_adp.png}
\caption{\label{fig:Mo_eam_adp} $\rho(r)$, $F(\rho)$, $\phi(r)$ of the original 
Zjw04 EAM, ML-EAM and ML-ADP. \textbf{(d)} shows the $\mu(r)$ and $\omega(r)$ of 
ML-ADP.}
\end{figure*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Mo - MAE
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lcccc}
\hline
 MAE   & SNAP & EAM & ML-EAM & ML-ADP \\
\hline
Energy (meV/atom) & 13.2 & 58.9 & 21.7 & 22.8 \\
Force (eV/\AA) & 0.25 & 0.31 & 0.28 & 0.27 \\
\hline
\end{tabular}
\caption{\label{table:elementary_Mo_mae} Comparion of the energy and force 
accuracy of the elementary SNAP Ni model, the original EAM, the MEAM and our 
ML-EAM.
}
\end{table}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Mo - Material Properties
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}[htp]
\centering
\begin{tabular}{lcccccc}
\hline
    & DFT & SNAP & EAM & ML-EAM & ML-ADP & Exp. \\
\hline
$T_{m}$(K) & & 3000 & 3750 & 2940 & 2880 & 2890 \\
$c_{11}$ (GPa) & 472 & 473 (0.2\%) & 456 (-3.4\%) & 260 (-5.8\%) & 274 (-0.7\%) & 479 \\
$c_{12}$ (GPa) & 158 & 152 (-3.8\%) & 167 (5.7\%) & 151 (0.0\%) & 163 (2.5\%) & 165 \\
$c_{44}$ (GPa) & 106 & 107 (0.9\%) & 115 (8.5\%) & 131 (0.0\%) & 131 (-0.8\%) & 108 \\
$E_{v}$ (eV) & 2.87 & 2.61 (-9.1\%) & 3.02 (5.2\%) & 1.16 (-20.5\%) & 1.71 (17.1\%) & \\
$E_{m}$ (eV) & 1.12 & 1.39 (24.1\%) & 1.54 (37.5\%) & 1.46 (30.5\%) & 0.87 (-22.3\%) & \\
$E_{a} = E_{v} + E_{m} $ (eV) & 3.99 & 4.00 (-0.1\%) & 4.56 (14.3\%) & 2.62 (1.6\%) & 2.58 (0.0\%) & 4.00 \\
\hline
\end{tabular}
\caption{\label{table:elementary_Mo} Comparison of the calculated and 
experimental material properties of elementary bcc Mo. SNAP refers to the 
elementary Mo SNAP model.
}
\end{table*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.C
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Mo-Ni}
\label{sec:alloy}

@@

The above results suggest that with machine learning and essential physical 
constraints, for the fcc and bcc solids, the traditional empirical potentials, 
EAM/ADP, can be improved to be as accurate as the SNAP method. However, 
EAM/ADP are around $10^2$ to $10^3$ times faster than SNAP.

The many-body expansion (MBE) is one of the most widely used schemes for fitting 
potential energy surfaces. In the many-body expansion scheme, the total energy 
of a system with N atoms can be expressed as the sum of all k-body terms where 
$k \le 1$:
\begin{equation}
\label{eq:many_body_expansion}
E = 
\sum_{i}^{\mathrm{C^N_1}}{E^{(1)}_{i}} +
\sum_{i,j}^{\mathrm{C^N_2}}{E^{(2)}_{ij}} + 
\sum_{i,j,k}^{\mathrm{C^N_3}}{E^{(3)}_{ijk}} + \cdots 
\end{equation}
where $\mathrm{C^N_k}$ is the binomial coefficient and $E^{(k)}$ represents the 
k-body contribution. In many cases, higher-order terms like $E_{ij}$ or 
$E_{ijk}$ are symmetric: $E_{ij}=E_{ji}$, $E_{ijk}=E_{ikj}=E_{jik}=\cdots$. 
Thus, Equation \ref{eq:many_body_expansion} can be further transformed to:
\begin{align}
\label{eq:MBE_atomic}
E & = \sum_{i}^{\mathrm{N}}{\left(
    E^{(1)}_{i} + 
    \frac{1}{2!}\sum_{j \ne i}^{\mathrm{N}}{E^{(2)}_{ij}} +
    \frac{1}{3!}\sum_{j \ne i}^{\mathrm{N}}{
        \sum_{k \ne i,j}^{\mathrm{N}}{E^{(3)}_{ijk}}
    } +
    \cdots
\right)} \nonumber \\
& = \sum_{i}^{\mathrm{N}}{E_{i}}
\end{align}
where $E_{i}$ is the atomic energy of atom $i$.

In fact, the EAM formalism can be considered as a variant of MBE truncated at 
$k=2$ (Equation \ref{eq:eam_total_energy}). The embedding contribution $F(\rho)$ 
serves as the one-body term. The pairwise contribution 
$\frac{1}{2}\sum_{j\ne i}{\phi(r_{ij})}$ is just a plain transcription of the 
two-body term in Equation \ref{eq:MBE_atomic}. The ADP formalism uses a much 
more complicated two-body term by introducing the pairwise dipole and quadrupole 
terms.

Although the SNAP method is quite complicated, the current formalism can still 
be viewed as a variant of MBE truncated at $k=2$. The atomic energy
